\Large{Adversarial Training against Systematic Uncertainty}

### 2016-10-27 tex/paper.tex
fix ref

In collisions at the Large Hadron Collider, massive particles can be produced at
such high energies that their decay products are collimated and the resulting
jets overlap. Still, deducing whether an observed jet is either due  to a single
hadronic particle or to the decay of a massive object into multiple hadronic
particles  is critical for understanding the nature of the particles produced in
the collision \citep{butterworth2008jet,guest2016jet}. Critically, this
classification problem is often made difficult by the presence of simultaneous
interactions, known as pileup, which contributes significant energy depositions
unrelated to the particles of interest.

In this setting, we reuse the classification problem of \citep{guest2016jet}
between single jets produced in quark or gluon fragmentation and two overlapping
jets produced when a high-velocity $W$ boson decays to a collimated pair of
quarks. As nuisance parameter, we consider events without pileup ($Z=0$) and
events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
50$ unrelated interactions are overlaid. Our goal is to build an accurate
classifier, for which we also want to minimize the effects due to the
uncertainties on the nuisance. More specifically, we choose to recast the
classification problem as a hypothesis test between signal+background (jets
originating from $W$ bosons or from single quarks and gluons) and background
only (jets originating from single quarks and gluons only) and tune the decision
threshold of our classifier by maximizing its approximate median significance
(AMS), when uncertainties in the background are taken into account (see Eqn.~20
of \citep{adam2014higgs}). Our motivation is that reducing the effects of
uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
allow for a larger maximum significance.

To minimize the effects of $Z$ in the background events, we train a classifier
using  Algorithm~\ref{alg:adversarial-training} but consider the adversarial
term ${\cal L}_r$ conditioned on $Y=0$ only. Both $f$ and $r$ are neural
networks with 3 hidden layers of 64 nodes and ReLU activations, each terminated
by a single final output node with a sigmoid activation. Experiments are performed
on the high-level features data described in \citep{guest2016jet}, on a subset
of 150000 samples for training while AMS is evaluated on an independent test set of 5000000
samples. Results reported below are averages over 5 runs.

>>>>

In collisions at the Large Hadron Collider, massive particles can be produced at
such high energies that their decay products are collimated and the resulting
jets overlap. Still, deducing whether an observed jet is either due  to a single
hadronic particle or to the decay of a massive object into multiple hadronic
particles  is critical for understanding the nature of the particles produced in
the collision \citep{butterworth2008jet,baldi2016jet}. Critically, this
classification problem is often made difficult by the presence of simultaneous
interactions, known as pileup, which contributes significant energy depositions
unrelated to the particles of interest.

In this setting, we reuse the classification problem of \citep{baldi2016jet}
between single jets produced in quark or gluon fragmentation and two overlapping
jets produced when a high-velocity $W$ boson decays to a collimated pair of
quarks. As nuisance parameter, we consider events without pileup ($Z=0$) and
events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
50$ unrelated interactions are overlaid. Our goal is to build an accurate
classifier, for which we also want to minimize the effects due to the
uncertainties on the nuisance. More specifically, we choose to recast the
classification problem as a hypothesis test between signal+background (jets
originating from $W$ bosons or from single quarks and gluons) and background
only (jets originating from single quarks and gluons only) and tune the decision
threshold of our classifier by maximizing its approximate median significance
(AMS), when uncertainties in the background are taken into account (see Eqn.~20
of \citep{adam2014higgs}). Our motivation is that reducing the effects of
uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
allow for a larger maximum significance.

To minimize the effects of $Z$ in the background events, we train a classifier
using  Algorithm~\ref{alg:adversarial-training} but consider the adversarial
term ${\cal L}_r$ conditioned on $Y=0$ only. Both $f$ and $r$ are neural
networks with 3 hidden layers of 64 nodes and ReLU activations, each terminated
by a single final output node with a sigmoid activation. Experiments are performed
on the high-level features data described in \citep{baldi2016jet}, on a subset
of 150000 samples for training while AMS is evaluated on an independent test set of 5000000
samples. Results reported below are averages over 5 runs.
